---
title: "Fun with Lags"
author: "David J Jackson"
date: "6/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RSQLite)
```
```{r, echo=FALSE}
rm(list=ls())
db <- dbConnect(SQLite(), dbname="../COVIDDB/CORVID.sqlite3")

us_states <- dbGetQuery(db, "SELECT state,date,new_cases,new_deaths FROM STATESDAILY;")
```
```{r, echo=FALSE}
ST <- us_states%>% filter(date >="2020-06-22") 

df <- ST %>% pivot_wider(
    names_from=date,
    values_from=starts_with('new_'))
```

### States Reporting a increase in Cases and Deaths for June 24, 2020

```{r, echo=FALSE}

df$Cases <- df$`new_cases_2020-06-24` -df$`new_cases_2020-06-23`
df1 <- df %>% filter( Cases >0)

df1 %>% ggplot(aes(state,Cases)) +geom_col() +
  coord_flip() + labs(title="States that had a increase in Cases",
                      subtitle ="From  June 23,2020 - June 24, 2020")


df$Deaths <- df$`new_deaths_2020-06-24` -df$`new_deaths_2020-06-23`
df2 <- df %>% filter( Deaths >0)

df2 %>% ggplot(aes(state,Deaths)) +geom_col() +
  coord_flip() + labs(title="States that had a increase in Deaths",
                      subtitle ="From  June 23,2020 - June 24, 2020")
```

### States Reporting Increase in Cases and Deaths for June 23,2020

```{r, echo=FALSE}
df$Cases1 <- df$`new_cases_2020-06-23` -df$`new_cases_2020-06-22`
df3 <- df %>% filter( Cases1 >0)

df3 %>% ggplot(aes(state,Cases1)) +geom_col() +
  coord_flip() + labs(title="States that had a increase in Cases",
                      subtitle ="From  June 22,2020 - June 23, 2020")


df$Deaths1 <- df$`new_deaths_2020-06-23` -df$`new_deaths_2020-06-22`
df4 <- df %>% filter( Deaths1 >0)

df4 %>% ggplot(aes(state,Deaths1)) +geom_col() +
  coord_flip() + labs(title="States that had a increase in Deaths",
                      subtitle ="From  June 22,2020 - June 23, 2020")

