---
title: "Ohio COVID-19 Daily Report"
author: "David Jackson"
date: "03/21/2021 2:25 PM/EDT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,error = FALSE)
library(ggplot2)
library(scales)
library(RSQLite)
library(dplyr)
library(forecast)
theme_set(theme_linedraw() + theme(panel.grid = element_line(linetype = 'dashed')))
```

```{r}
### State Data
rm(list=ls())
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df <- dbGetQuery(db,"select * from STATESDAILY")
df <- df[order(df$date),]
df <- subset(df,state =="Ohio" & date >="2020-03-10" )
df$date <- as.Date(df$date)

sum(df$new_cases,na.rm=TRUE)
sum(df$new_deaths,na.rm=TRUE)
## County Data
DF <- dbGetQuery(db,"select date,state,county,new_cases,new_deaths from  COUNTYDAILY where state='Ohio'")
DF$date <- as.Date(DF$date)
str(DF$date)
```

## Big New's About Ohio's Covid-19 Deaths.(2021-02-12)
* "Ohio health dept. to restructure after 4K unreported deaths"
* Source: https://wpta21.com/2021/02/12/governor-expected-to-discuss-underreported-covid-19-deaths/
* Summary: Ohio "un·der·re·ported" Covid-19 Deaths  by as much  4,000
* Thursday’s daily coronavirus death toll showed more than 720 deaths, of which 650 come from previously unreported deaths, GOP Gov. Mike DeWine said.(2021-02-11)
* "The Health Department said that “process issues affecting the reconciliation and reporting of these deaths” began in October, with most occurring in November and December"

#### Mortality Rate by Date

```{r}
ggplot(df) + geom_line(aes(x=date,y=death_rate)) +
  scale_y_continuous(labels=percent) +
  geom_hline(yintercept = 0.025,col="red") +
  ylim(0,0.3)

```

#### Daily COVID19 Cases and Deaths

```{r}

US <- df %>% filter(date >="2020-04-01")
Trend <- US
US$MAC <- ma(US$new_cases,14,centre=TRUE)
US$MAD <- ma(US$new_deaths,14,centre=TRUE)

ggplot(US,aes(x=date,y=MAC)) + geom_col() +
    labs(title="Ohio Daily COVID19 Cases by Date",
         subtitle = "(14 Day Moving Average)") 
  

ggplot(US,aes(x=date,y=MAD)) + geom_col() +
    labs(title="Ohio Daily COVID19 Deaths by Date",
         subtitle = "(14 Day Moving Average/ y-limit:0,100)") +
  ylim(0,100)
```

#### Weekly and Monthly Cases and Deaths

```{r}
US$Monthly <- as.Date(cut(US$date,
  breaks = "month"))
US$Weekly <- as.Date(cut(US$date,
  breaks = "week",
  start.on.monday = FALSE))
```
```{r}
Weekly_new_cases <- aggregate(new_cases~Weekly,US,FUN=sum)
Weekly_new_deaths <- aggregate(new_deaths~Weekly,US,FUN=sum)

```
```{r}
ggplot(Weekly_new_cases,aes(x=Weekly,y=new_cases)) + geom_col() +
  geom_smooth(span=0.25,se=FALSE) + labs(title="Weekly COVID19 Cases")

ggplot(Weekly_new_deaths,aes(x=Weekly,y=new_deaths)) + geom_col() +
  geom_smooth(span=0.25,se=FALSE) + labs(title="Weekly COVID19 Deaths") +
  ylim(0,1000)
```
```{r}
monthly_new_cases <- aggregate(new_cases~Monthly,US,FUN=sum)
monthly_new_deaths <- aggregate(new_deaths~Monthly,US,FUN=sum)
```
```{r}
ggplot(monthly_new_cases,aes(x=Monthly,y=new_cases)) + geom_col() +
  labs(title="Monthly COVID19 Cases")

ggplot(monthly_new_deaths,aes(x=Monthly,y=new_deaths)) + geom_col() +
   labs(title="Monthly COVID19 Deaths")
```

### Top Five Counties Based on Daily Totals for Previous 8 Days

```{r}
df <- subset(DF,date ==Sys.Date() -8 )
Top <- df[order(df$new_cases,decreasing = TRUE),]
Top5 <-head(Top,n=5)
ggplot(Top5,aes(x=new_cases,y=county)) +
  geom_col() + coord_flip() +
  labs(title="Top 5 Counties by Total Cases for last 7 day.")

Top5 <-head(Top,n=5)
ggplot(Top5,aes(x=new_deaths,y=county)) +
  geom_col() + coord_flip() +
  labs(title="Top 5 Counties by Total Deaths for last 7 day.")
```

### Plot of Franklin Country Daily Cases and Deaths

```{r}
franklin <- subset(DF,DF$county =="Franklin")
ggplot(franklin,aes(x=date,y=new_cases)) + geom_line() +
  geom_smooth(span=0.25) +
  labs(title="Franklin County Cases by Date")

franklin <- subset(DF,DF$county =="Franklin")
ggplot(franklin,aes(x=date,y=new_deaths)) + geom_line() +
  geom_smooth(span=0.25) +
  labs(title="Franklin County Deaths by Date") + ylim(0,20)
```

#### Fixed 4th quarter numbers (do to under counting of deaths)

```{r}
# df$Ajusted <-  ifelse(df$date >="2020-10-01" & df$date <="2020-12-31",df$new_deaths + 40,df$new_deaths)
# 
# ggplot(df) + geom_line(aes(x=date,y=new_deaths,col="Deaths")) +
#   geom_line(aes(x=date,y=Ajusted,col="Adjusted")) +
#   ylim(0,200)
```

